---
title: "Investigating Indie"
output: 
  flexdashboard::flex_dashboard
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(ggplot2)
library(spotifyr)
library(compmus)
library(plotly)
library(gridExtra)

folk <- get_playlist_audio_features("", "6P82OCSEpo1VrZ4TZ2W5jK")
pop <- get_playlist_audio_features("", "3HDdTtyB76aLDc4wpGFCX4")
rock <- get_playlist_audio_features("","5dMoFfrgm8aOOkneQjt6LT")

indie <-
  bind_rows(
    folk |> mutate(category = "Indie Folk"),
    pop |> mutate(category = "Indie Pop"),
    rock |> mutate(category= "Indie Rock")
  )
```

Introduction
===================================== 

### What makes a song "Indie"? 
  There is an ongoing debate on whether indie music can be considered a genre. Indie is short for “independent” and refers to music released by an artist independently rather than with a record label. Some bands that are considered to be indie, such as Arcade Fire, however, work with a record label. This begs the question: What makes indie music indie? Some argue that indie music has a specific sound, but what are its defining characteristics that allows Spotify to create “Indie” playlists? I hope to investigate this question with my corpus containing songs from three different indie sub-genre playlists created by Spotify: “Indie Rock Mix,” “Indie Pop Mix,” and “Indie Folk Mix.” Comparing songs from these different sub-genres will clarify what characteristics of a song Spotify considers to be indie. The main limitation of my investigation is that the songs in the playlists are recommended by Spotify for my specific account. Hence, it may not reflect the most popular indie songs nor be representative of indie music as a whole. “The Less I Know the Better” by Tame Impala is a typical indie song in my corpus, because it cannot be tied to another genre. Two atypical songs are “Sunflower” by Harry Styles and “Work Song” by Hozier, because neither of these artists are considered to be indie artists. These songs will be good to analyze since they may have the characteristics of a typical indie song, but they are made by non-indie artists.
  
  
```{r}

```


Comparing the Three Playlists {.storyboard}
===========================================================

### Indie Keys
```{r}
key_histogram <- ggplot(indie, aes(x=key, color=category,fill=category)) + geom_histogram(binwidth = 1, alpha=0.6) + scale_x_continuous(breaks = c(0:11), labels=c("C","C#","D","D#","E","F","F#","G","G#","A","A#","B")) + 
  facet_wrap(~category) + labs(y = "Count", x="Key", title = "Comparing the Distribution of Key/nBetween Indie Subgenres") + theme(legend.title = element_blank())
ggplotly(key_histogram)
```
***
DESCRIPTION


### Indie Rock is Energetic
```{r}
dens_energy <- ggplot(indie, aes(x = energy, color = category)) + geom_histogram(fill="white") + facet_wrap(~category) + 
  labs(x = "Energy", y = "Density", 
       title = "Comparing the Distribution of Energy Between Indie Subgenres",
       color = "Indie Subgenres") 
ggplotly(dens_energy)
```

*** 
This histogram shows that energy is widespread over the Indie genre as a whole. A song's energy cannot tell us whether it belongs under the "Indie" label. What is clear, however, is that Spotify's "Indie Rock" playlist holds the songs with the most energy. Hence, Indie Rock separates itself from the other two subgenres through Spotify's energy audio feature.

### Indie Rock Has The Energy, Indie Folk Is For The Feels, Indie Pop Is Undecided {data-commentary-width=400}
```{r}
energy_valence <- ggplot(indie, aes(x = energy, y = loudness, color=valence, label = track.name)) + geom_point() + facet_wrap(~category) + scale_color_gradient(low = "yellow", high = "purple") + labs(x = "Energy", y = "Loudness") + geom_smooth(method=lm)
ggplotly(energy_valence)
```

***

Consistent with the last histogram, energy and loudness are widespread over the Indie genre as a whole. What can be observed again is that Spotify's "Indie Rock" playlist holds the songs with the most energy and loudness, followed by "Indie Pop" and "Indie Folk." Hence, the energy and loudness features can help us distinguish Indie songs from each other and categorize them into sub-genres. 
Indie Folk separates itself from the other two sub-genres through the valence feature. Compared to Rock and Pop, most of the Indie Folk songs have a negative valence. In other words, Indie Folk is a sad sub-genre.
The song "j's lullaby" by Delaney Bailey is a clear outlier in this graph. It has the least energy and loudness by far compared to not only the other indie pop songs but also compared to the songs in the other two sub-genres. Another atypical song is "Naked as We Came" by Iron & Wine since it is one of the few songs that is low in energy and loudness but has a positive valence. These songs will be further investigated.


Comparing Pitch {data-navmenu="A Closer Look"}
==================================

Column {data-width=700}
-----------------------------------
### Energetic Indie Has A More Colorful Pitch Than Quiet Indie
```{r}
j_lullaby <-
  get_tidy_audio_analysis("7LNm1oWd6PniO9wRYpNY5o") |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches)

j_lullaby_plot <- j_lullaby |>
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
  compmus_gather_chroma() |>
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = pitch_class,
      fill = value
    )
  ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude", title = "\"j's lullaby\" by Delaney Bailey", subtitle = "Indie Pop, Lowest Energy, Low Valence") +
  theme_minimal() +
  scale_fill_viridis_c() + theme(plot.title = element_text(face="bold", size=8), plot.subtitle = element_text(size=6), axis.title = element_text(size=6), axis.text = element_text(size=6) , legend.title = element_text(size=6), legend.key.size = unit(0.3, 'cm'), legend.text = element_text(size=5))

live_once <-
  get_tidy_audio_analysis("4fPBB44eDH71YohayI4eKV") |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches)

live_once_plot <- live_once |>
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
  compmus_gather_chroma() |>
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = pitch_class,
      fill = value
    )
  ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude", title = "\"You Only Live Once\" by The Strokes", subtitle = "Indie Rock, Highest Energy, High Valence") +
  theme_minimal() +
  scale_fill_viridis_c() + theme(plot.title = element_text(face="bold", size=8), plot.subtitle = element_text(size=6), axis.title = element_text(size=6), axis.text = element_text(size=6) , legend.title = element_text(size=6), legend.key.size = unit(0.3, 'cm'), legend.text = element_text(size=5))

naked <- get_tidy_audio_analysis("4FuBIkfhZMoRgrDiHL6TYG") |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches)

naked_plot <- naked |>
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
  compmus_gather_chroma() |>
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = pitch_class,
      fill = value
    )
  ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude", title = "\"Naked As We Came\" by Iron & Wine", subtitle = "Indie Folk, Low Energy, High Valence") +
  theme_minimal() +
  scale_fill_viridis_c() + theme(plot.title = element_text(face="bold", size=8), plot.subtitle = element_text(size=6), axis.title = element_text(size=6), axis.text = element_text(size=6) , legend.title = element_text(size=6), legend.key.size = unit(0.3, 'cm'), legend.text = element_text(size=5))

wolf <- get_tidy_audio_analysis("03wKMRNYVvw6s9nm4I4jUS") |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches)

wolf_plot <- wolf |>
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
  compmus_gather_chroma() |>
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = pitch_class,
      fill = value
    )
  ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude", title = "\"Wolf Like Me\" by TV On The Radio", subtitle = "Indie Rock, High Energy, Low Valence") +
  theme_minimal() +
  scale_fill_viridis_c() + theme(plot.title = element_text(face="bold", size=8), plot.subtitle = element_text(size=6), axis.title = element_text(size=6), axis.text = element_text(size=6) , legend.title = element_text(size=6), legend.key.size = unit(0.3, 'cm'), legend.text = element_text(size=5))

grid.arrange(j_lullaby_plot, live_once_plot, naked_plot, wolf_plot, ncol=2)
```

Column {data-width=300}
-----------------------------------
#### Chromagrams
To the left are four chromagrams of four different songs that are interesting to compare. A chromagram is a visual representation of the pitch of a piece of music. It shows how much energy is contained in each pitch class at any given moment. 

#### Analysis
Here we are comparing songs with different extreme levels of energy and valence. As discussed earlier, "j's lullaby" by Delaney Bailey is the song in the corpus with the lowest energy and valence. On the extreme opposite, "You Only Live Once" by The Strokes is one of the songs with the highest energy and highest valence. The other two songs displayed have an atypical combination of opposing energy and valences. The Indie Folk song "Naked As We Came" has low energy and high valence, while "Wolk Like Me" has high energy and low valence.

When comparing these four chromagrams, we can see that the two songs with high energy, "You Only Live Once" and "Wolf Like Me," have their signal's energy more widespread over all the pitch classes. "Wolf Like Me" has much of its energy contained in pitch class B, but the other pitch classes still have some energy (hints of yellow) displayed. "You Only Live Once" has much of its energy contained in pitch classes E and B, but again most of the energy is spread out across all the pitch classes. By contrast, the two songs with low energy have more distinct pitch classes. "j's lullaby" has most of it's energy focused on pitch classes C# and G#, and "Naked As We Come" has most of its energy contained in pitch classes E and B.


Self-Similarity Matrices {.storyboard data-navmenu="A Closer Look"}
==================================

### The Typical Indie Rock Song "A-Punk" by Vampire Weekend: Simple Chroma & Complex Timbre {data-commentary-width=400}
``` {r}
a_punk <-
  get_tidy_audio_analysis("3AydAydLzyyZutA0375XIz") |> # Change URI.
  compmus_align(bars, segments) |>                     # Change `bars`
  select(bars) |>                                      #   in all three
  unnest(bars) |>                                      #   of these lines.
  mutate(
    pitches =
      map(segments,
          compmus_summarise, pitches,
          method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  ) |>
  mutate(
    timbre =
      map(segments,
          compmus_summarise, timbre,
          method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  )

a_punk_chroma <- a_punk |>
  compmus_self_similarity(pitches, "cosine") |> # change timbre/pitches
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "", title = "Chroma-based Self-Similarity Matrix") + theme(title = element_text(size = 8))

a_punk_timbre <- a_punk |>
  compmus_self_similarity(timbre, "cosine") |> # change timbre/pitches
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "", title = "Timbre-Based Self-Similarity Matrix") + theme(title = element_text(size = 8))

grid.arrange(a_punk_chroma, a_punk_timbre, ncol=2)
```

***
A self-similarity shows the degree of similarity between different parts of a musical piece. It helps us understand how a song is structured through both its pitch (chroma) and timbre.

The self-similarity matrices on the right are those of a song by the Vampire Weekend. The Vampire Weekend is a known Indie Rock band. There debut album "Vampire Weekend" was independently released in 2008, hence officially falling under the "Indie" label. "A-Punk" is an energetic and youthful song that starts with a light-hearted guitar riff which is joined by drums and vocals.

It is short song with the following structure: verse, chorus, refrain, verse, chorus refrain. We can see this pattern in the chroma-based matrix: 0-40s is verse 1, 40-60s is chorus & refrain, 60-100s is verse 2, and then back to chorus and refrain. The chroma-based matrix helps us see this clear structure of the song. This pattern is most likely why the song has become so popular. The simple structure makes the song simple to follow and fun to dance to.

The timbre-based matrix is more interesting to look at. At 40 seconds, the first change in timbre is marked by the flute entering the song, which gives the song an extra layer of timbre complexity. At 60 seconds, the cymbal is abruptly played alone, followed by a vocal "hey hey hey" with no instrumental background, and then the drums come back in to move onto the next verse. This is the part of the song that gives the listener anticipation for what is coming next. After the second verse, the flute comes back followed by the "hey hey hey." At the end of the song (~130s), the cymbal comes back abruptly alone marking the end of the song.

This typical indie rock song shows that the changes pitches and vocals are not as important as the changes in timbre in indie rock. The instruments of the bands play an important role in creating a successful indie rock song.


### Typical Indie Folk: Simple Chroma & Simple Timbre {data-commentary-width=400}
``` {r}
flightless <-
  get_tidy_audio_analysis("3AydAydLzyyZutA0375XIz") |> # Change URI.
  compmus_align(bars, segments) |>                     # Change `bars`
  select(bars) |>                                      #   in all three
  unnest(bars) |>                                      #   of these lines.
  mutate(
    pitches =
      map(segments,
          compmus_summarise, pitches,
          method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  ) |>
  mutate(
    timbre =
      map(segments,
          compmus_summarise, timbre,
          method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  )

flightless_chroma <- flightless |>
  compmus_self_similarity(pitches, "cosine") |> # change timbre/pitches
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "", title = "Chroma-based Self-Similarity Matrix") + theme(title = element_text(size = 8))

flightless_timbre <- flightless |>
  compmus_self_similarity(timbre, "cosine") |> # change timbre/pitches
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "", title = "Timbre-Based Self-Similarity Matrix") + theme(title = element_text(size = 8))

grid.arrange(top="\"Flightless Bird, American Mouth\" by Iron & Wine", flightless_chroma, flightless_timbre, ncol=2)
```